---
title: "Guided Project : Exploring NYC Schools Data"
author: Nupur Gupta
date: 7/22/2019
output: html_notebook
---
```{r}
library(dplyr)
library(readr)
library(purrr)
library(ggplot2)
library(stringr)
install.packages("tidyr")
library(tidyr)
```


```{r}
combined <- read_csv("combined.csv")
gened <- read_tsv("masterfile11_gened_final.txt")
d75 <- read_tsv("masterfile11_d75_final.txt")
```

Responses to surveys designed to gauge parent, student, and teacher perceptions of the quality of New York City schools. In this guided project, we'll guide you through investigating the following questions:

Do student, teacher, and parent perceptions of NYC school quality appear to be related to demographic and academic success metrics?

Do students, teachers, and parents have similar perceptions of NYC school quality?

Data Sources: https://data.cityofnewyork.us/Education/2011-NYC-School-Survey/mnz3-dyi8
"combined" data is downloaded from dataquest.io.

Since we are interested in what we can learn from survey data about perceptions of school quality to study its relationship with SAT scores, we can likely focus our efforts on the aggregate scores for each school calculated from the survey responses.

```{r}
gened <- gened %>%
  select(dbn, schoolname, schooltype, rr_s:aca_tot_11) %>%
  filter(schooltype == "High School")

d75 <- d75 %>%
  select(dbn, schoolname, rr_s:aca_tot_11) 


survey <- bind_rows(d75, gened)
survey <- survey %>%
  rename(DBN = dbn)

combined <- combined %>%
  left_join(survey, by = "DBN")

```

Used left join to combine the three data frames.

```{r}
cor_mat <- combined %>%
  select(avg_sat_score, saf_p_11:aca_tot_11) %>%
  cor(use = "pairwise.complete.obs")

cor_tib <- cor_mat %>%
  as_tibble(rownames = "variables")

strong_cor <- cor_tib %>%
  select(variables, avg_sat_score) %>%
  filter(avg_sat_score > 0.25 | avg_sat_score < -0.25)

scatterfun <- function(x,y) {
  ggplot(data = strong_cor) +
    aes(x = x, y = y) + 
    geom_point(alpha = 0.3)
}

x_var <- strong_cor$variables[2:5]

map2(x_var, "avg_sat_score", scatterfun)

```


```{r}
combined_gather <- combined %>%
  gather(key = "survey_ques", value = score, saf_p_11:aca_tot_11)

combined_gather <- combined_gather %>%
  mutate(response_type = str_sub(survey_ques, 4,6)) %>%
  mutate(metric = str_sub(survey_ques,1,3))

combined_2 <- combined_gather %>%
  mutate(response_type = ifelse(response_type == "_p_", "parent", ifelse(response_type == "_s_", "student", ifelse(response_type == "_t_", "teacher", ifelse(response_type == "_to", "total", "NA"))))) %>%
  mutate(metric = ifelse(metric == "saf", "safety and respect", ifelse(metric == "aca", "academic expectations", ifelse(metric == "eng", "engagement", "communication"))))


```

Some rough analysis:

```{r}
combined_2 <- combined_2 %>%
  filter(response_type != "total") 

ggplot(combined_2) +
  aes(x = metric, y = score, fill = response_type) +
  geom_boxplot()
```

Parents have a higher expectation/ score on an average compared to students and sometimes even to teachers. 
